### 2019-02-10 paper.tex
Wip

\documentclass{article}

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{hyperref}

\newcommand{\theHalgorithm}{\arabic{algorithm}}

\usepackage[accepted]{sysml2019}

\sysmltitlerunning{Guild AI}

\begin{document}

\twocolumn[
\sysmltitle{Guild AI: Simple Reproducibility in Machine Learning}

\sysmlsetsymbol{equal}{*}

\begin{sysmlauthorlist}
\sysmlauthor{Garrett Smith}{equal,gai}
\end{sysmlauthorlist}

\sysmlaffiliation{gai}{Garrett Smith, Chicago, USA}

\sysmlcorrespondingauthor{Garrett Smith}{garrett@guild.ai}

\sysmlkeywords{Machine Learning, SysML, Guild AI}

\vskip 0.3in

\begin{abstract}
  Guild AI is an open souce tool for running, tracking, and comparing
  machine learning experiments. Guild helps researchers and engineers
  do their work more efficiently without imposing additional
  configuration and system requirements. By simplifying experiment
  tracking, Guild hopes to encourage consistent reproducibility and
  improved collaboration in machine learning research.
\end{abstract}
]

\printAffiliationsAndNotice{\sysmlEqualContribution}

\section{Guild AI}

Guild AI is a command line tool that runs researcher code without
modification. It automatically captures results as unique experiments
and provides facilities to analyze, compare, and share them.

\textbf{Experiments.} Guild runs experiment trials as operating system
processes and tracks results in separate run directories. Each run
directory contains the artifcats generated during the trial along with
metadata such as model information, start and stop times, logs, and
run status. Run artifacts commonly include training checkpoints,
prepared data sets, generated content such as images and audio, logs,
and project source code snapshots. Guild does not require or assume
any particular generated output from trials.

\textbf{Analysis.} Guild provides facilities to analyze and compare
experiments. User can study experiment outputs such as training and
validation loss, accuracy, required memory, batch latency---any metric
recorded across experiments---as well logs and generated files. Guild
also supports detailed comparison across experiment inputs including
changes in model architecture, hyperparameters, data sets, and source
code.

\textbf{Sharing Results.} Guild helps researchers share results in
three ways ways:

\begin{itemize}[topsep=3pt]
\item Users \emph{publish experiment results} to remote locations over
  SSH or cloud services like S3 where they can be studied and compare
  by others.
\item Project maintainers use \emph{Guild files}, which are simple,
  YAML files included with project source, to document and automate
  project capabilities, such as supported models, operations, default
  hyperparameter values and related search spaces for optimization
  experiments.
\item Project maintainers may further generated \emph{Python packages}
  that can be uploaded to PyPI and installed by others for reproducing
  results.
\end{itemize}

\textbf{Command Line Interface, Visualization, API.} Guild's primary
user interface is a command line interface, which integrates
efficiently with other system tools. Guild additionally provides a
graphical, web based dashboard to explore and analyze runs and
integration with TensorBoard. Guild exposes all core functionality in
a Python API for application developers.

\textbf{Cross Framework and Language Support.} Guild supports any
learning script that generates experiment results. Models can be
implemented in any language, framework, or software library. Commonly
supported libraries include TensorFlow, Keras, PyTorch, MXNet,
scikit-learn, and SGBoost. However, Guild supports other Python
libraries, pure Python, and languages such as R, Go, and Java.

\subsection{Reproducibility}

Guild AI facilitates reproducibility across trials by capturing
results and providing tools for comparison. A typical workflow in
Guild is:

\begin{enumerate}[noitemsep, topsep=0pt, partopsep=0pt]
\item Run trials for a baseline model
\item Run trials for novel work
\item Compare and analyze results
\end{enumerate}

Additionally, Guild supports packaging work for distribution so others
can quickly and easily install and run trials, which in turn establish
baselines for their work.

\subsection{Users}

Guild is used by researchers and research engineers to run, track, and
compare machine learning experiments. It is useful to anyone who
outinely runs experiments that compare novel work to baseline results.

\subsection{Related Work and Differences}

There are several open source projects that address reproducibility in
machine learning. Prominent projects include:

\begin{itemize}[noitemsep, topsep=0pt, partopsep=0pt]
\item ModelDB \cite{modeldb}
\item MLFlow \cite{mlflow}
\item Polyaxon \cite{polyaxon}
\item datmo \cite{datmo}
\end{itemize}

Automation tools in machine learning tend to require tool-specific
changes to project source code. This generally includes adoption of a
Python API for controlling experiments, accessing hyperparameters and
logging results. Furthermore, automation tools often require
additional system software such as databases, container management
systems, and job schedulers.

Each requirement that a tool imposes on a user adds friction to the
process of experiment automation and tracking. Researchers may
perceive the cost of reproducibility to be greater than the benefits.

In contrast, Guild does not require changes to user code or
installation and maintenance of external systems. Users can run their
existing training scripts without modification. For example, an
unmodified Python script \texttt{train.py} that supports
hyperparameters \texttt{n\_layers} and \texttt{lr} may be run using:

{\footnotesize
\begin{verbatim}
$ guild run train.py n_layers=2 lr=1e-4
\end{verbatim}}

With this command, the user generates a unique experiment with the
specified hyperparameters that can be analyzed and compared to other
trials. Guild manages the entire workflow without external
dependencies.

Guild additionally supports easy packaging, distribution, and
installation of research work in the support of reproducibility. To
our knowledge, no other tool provides this feature.

\section{Demonstration}

We propose an interactive scenario that highlights the benefits of
fast, easy experiment reproduction. During the demo, he presenter
refers to two personas, which are represented visually with placards,
to help viewers maintain context as stages of the scenario are
presented:

\begin{itemize}[topsep=3pt]
\item \emph{Arya}. A researcher who has published state-of-the-art
  results of a novel architecture on a standard benchmark dataset.
\item \emph{Clegane}. A researcher working on a survey paper that
  includes Arya's work.
\end{itemize}

While demonstrating a scenario, the presenter will highlight how the
tool supports reproducibility in two areas:

\begin{itemize}[topsep=3pt]
\item \emph{Influence}. Work that is easier to reproduce is more
  likely influence others, other factors held constant.

\item \emph{Productivity}. Facilities that simplify and speed
  reproducibility are applied equally to trial generation in general.
  Researchers can learn from more experiments in less time to improve
  results.
\end{itemize}

\textbf{Scenarios}. The demonstration is structured flexibly to
encourage questions, input, and redirection from audience
participants. While presented below as a natural sequence, scenarios
may be presented and discussed in any order as needed to meet audience
expectation and interest.

\begin{itemize}[topsep=3pt]
\item \emph{Arya} uses Guild to run training and validation
  experiments, collecting results---both positive and negative---and
  announces her findings in a celebrated invited talk at SysML. Her
  work to support reproducibility is done when she commits her code to
  GitHub.

\item \emph{Clegane} conducts an survey of related work, reproducing
  published results where possible. He clones Arya's repository and
  uses Guild to reproduce her experiments within his environment. He
  has time to experiment with changes to Arya's model and default
  hyperparameters to build a stronger intuition for his summary.

\item \emph{Clegane} contacts Arya to discuss her work in more detail
  and answer some questions he has based on his own experiments. He
  uses Guild to publish experiments to S3 where Arya can access them.

\item \emph{Arya} pulls Clegane's results from S3 and uses Guild to
  study the differences between her experiments and Clegane's. She
  uses this information to answer Clegane's questions, which he
  includes in his paper.

\item \emph{Clegane}, to promote his field of study and encourage
  further research, creates an installable package that anyone can use
  to reproduce Arya's work.
\end{itemize}

\subsection{Equipment}

The presenter supplies a laptop with HDMI and brings supporting
graphical placards and other visual aids to reference during
scenarios. To avoid problems with network connectivity, the
demonstration is disconnected from the network. Demonstrations of
networking features (e.g. copying experiment results to and from S3)
are run locally over a proxy, but are nonetheless live and unmocked.

\bibliography{paper}
\bibliographystyle{sysml2019}

\end{document}